---
title: "pml"
author: "Alejandro"
date: "23 de agosto de 2015"
output: html_document
---



I started getting the data and making some data partition.
I used 80% of ther data to train the model

```{r cache=TRUE}
setwd("D:/data/Desktop/PracticalMachineLearning/pml")
set.seed(178632)
training <- read.csv("pml-training.csv")


test <- read.csv("pml-testing.csv")
library(caret)

inTrain<-createDataPartition(y=training$classe,p=.80,list = FALSE)

summary(inTrain)
trainingSample<-training[inTrain,]
testingSample<-training[-inTrain,]
summary(trainingSample)

```
After this rpocess was made I remove the low variance variables
The columns were removed from training and testing datasets
```{r cache=TRUE}
nzv <- nearZeroVar(trainingSample, saveMetrics= TRUE)
length(nzv[which(nzv$nzv==TRUE),1])

nzv <- nearZeroVar(trainingSample)
#data without zero vars
trainingSampleCLean <- trainingSample[, -nzv]

#Appling to testing
testingSample <- testingSample[, -nzv]
```

Adicionaly I remove the first 6 columns which does not get any value to the model

```{r cache=TRUE}

trainingSampleCLean<-trainingSampleCLean[,-(1:6)]

#Appling to testing
testingSample <- testingSample[, -(1:6)]

```
In the trainig dataset summary I realized that there are many variables with NAs, these variables have to be remove in order to get better results

```{r cache=TRUE}
selcols = grep("^max_|^min_|^amplitude_|^var_|^avg_|^stddev_", names(trainingSampleCLean)) 
summary(trainingSampleCLean[,-selcols] )

trainingSampleCLean<-trainingSampleCLean[,-selcols] 
#Appling to testing
testingSample <- testingSample[, -selcols]

```
Now I  have a better dataset, but it's necessary to clean the data a little bit more and remove columns that are higtly correlated

```{r cache=TRUE}

descrCor <-  cor(trainingSampleCLean[,1:length(trainingSampleCLean)-1])

highlyCorDescr <- findCorrelation(descrCor, cutoff = .75)

trainingSampleCLean <- trainingSampleCLean[,-highlyCorDescr]

#Appling to testing
testingSample <- testingSample[, -highlyCorDescr]

```
Now we have a dataset with varibles correlated no mare than 0.75

```{r cache=TRUE}
descrCor2 <- cor(trainingSampleCLean[,1:length(trainingSampleCLean)-1])
summary(descrCor2[upper.tri(descrCor2)])

```
Now I start with the model fitting
First I set the parameters for the training, after that I train two models and rpart model and the random forest.

```{r cache=TRUE}
set.seed(1745)
folds=10
repeats=10
fitControl <- trainControl(method="cv",number=folds,repeats=repeats,
                           classProbs=T,
                           allowParallel=T
                           )

modfit1 <- train(classe~., data=trainingSampleCLean,
                     method="rpart",metric="ROC",
                     trControl=fitControl
                     )






predictedTrain1<-predict(modfit1,newdata = trainingSampleCLean)

predictedTest1<-predict(modfit1,newdata = testingSample)





confusionMatrix(predictedTrain1,trainingSampleCLean$classe)
confusionMatrix(predictedTest1,testingSample$classe)



```
```{r cache=TRUE}
modfit2<-train(classe~.,method="rf", trControl=trainControl(method='cv'), number=5, allowParallel=TRUE ,data=trainingSampleCLean)
predictedTrain2<-predict(modfit2,newdata = trainingSampleCLean)

predictedTest2<-predict(modfit2,newdata = testingSample)
confusionMatrix(predictedTrain2,trainingSampleCLean$classe)
confusionMatrix(predictedTest2,testingSample$classe)


```



Applying the model to the test set


```{r cache=TRUE,eval=FALSE}
test <- test[, -nzv]
test <- test[, -(1:6)]
test <- test[, -selcols]
test <- test[, -highlyCorDescr]
answers<-predict(modfit2,newdata = test)

setwd("D:/data/Desktop/PracticalMachineLearning/pml/resultR")
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_",i,".txt")
    write.table(x[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
  }
}

pml_write_files(answers)

```
